Learning constraints in spreadsheets and tabular data

### 2016-09-23 Paper/ECAI/tabular.tex
Added ordered constraints

  \caption{Constraints learned for the tables above, except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5) constraints.

@@

@@

    \eclookupfuzzy{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
      & \textit{Same as lookup}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.50\linewidth]{figures/constraint_dependency2.pdf}
  \caption{Dependency graph of Table~\ref{table:constraints}; an arrow from~$s_1$ to~$s_2$ indicates that $s_2$ depends on~$s_1$ (its signature includes $s_1$).
  }
  \label{fig:learning_order}
\end{figure}

@@

\subsubsection{Constraints} \label{sec:which_cons}
The \samuel{count} constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
We included most formulas that we encountered in tutorial spreadsheets including the popular \textit{SUM} and \textit{LOOKUP} constraints\footnote{\href{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}}.
We also added three structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION} and \textit{FOREIGNKEY}) so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.

@@

All spreadsheets also have a set of ground-truth of constraints, we call these the \textbf{intended} constraints, that are expected to be (re-)discovered.
These were determined by us using either the formulas of the original sheet (if present), or inferring them based on, for example, the header names in the sheet.
In the following experiments, we focus on spreadsheets specifically and hence only include \textit{functional} constraints, constraints that can be expressed as formulas. We ignoring structural constraints such as \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}. All constraints are stored in their canonical form. 5 spreadsheets contained rows or columns that were entirely identical, and any intended constraint involving one such vector was also added for the equivalent vectors. \samuel{They are usually not copied}

Section~\ref{sec:which_cons} discusses the 32 constraints currently supported by \sname \tias{It does, right? Should also contain '32' number}\sam{I will recount at the end}. Some of the intended constraints are currently outside of the scope of \sname however, more specifically nested mathematical or nested logical formulas. We denote by \textbf{supported} constraints the subset of intended constraints that the system can find in theory.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=1\linewidth]{figures/scatter_plots.pdf}
  \caption{Loglog plots of the experiments with synthetic (random) data. In plot~$a$ the number of vectors is increased; in plot~$b$ the size of the vectors is increased; and in plot~$c$ the number of blocks is increased (the block size is decreased).}
  \label{fig:runtime_analysis}
\end{figure}

@@

Our method uses dependencies to increase efficiency by finding constraints incrementally. To illustrate the effect this can have, we ran the benchmarks using FOREIGNKEY instead of ALLDIFFERENT as base constraint to find conditional aggregates such as SUMIF. This is a stronger assumption, and a too strong in practice as not all keys have to be used in the conditions. However, the total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data). This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.\sam{Check numbers}

@@

>>>>

  \caption{Constraints learned for the tables above, except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2), \textit{FOREIGNKEY} (5) and \textit{ORDERED} (6) constraints.

@@

    \ecordered{\sbl{x}}
      & $\numeric(\sbl{x})$

      & $i < j$: $\sbl{x}[i] \leq \sbl{x}[j]$.
      \\[\PADSEP] \hline

@@

    \eclookupfuzzy{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
      & \textit{Same as lookup}; except $\ecordered{\sbl{pk}}$ instead of \ecfkey{\sbl{fk}}{\sbl{pk}}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.60\linewidth]{figures/constraint_dependency.pdf}
  \caption{Dependency graph of Table~\ref{table:constraints}; an arrow from~$s_1$ to~$s_2$ indicates that $s_2$ depends on~$s_1$ (its signature includes $s_1$).
  }
  \label{fig:learning_order}
\end{figure}

@@

\subsubsection{Constraints} \label{sec:which_cons}
The 33 constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
We included most formulas that we encountered in tutorial spreadsheets including the popular \textit{SUM} and \textit{LOOKUP} constraints\footnote{\href{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}}.
We also added three structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION} and \textit{FOREIGNKEY}) so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.

@@

All spreadsheets also have a set of ground-truth of constraints, we call these the \textbf{intended} constraints, that are expected to be (re-)discovered.
These were determined by us using either the formulas of the original sheet (if present), or inferring them based on, for example, the header names in the sheet.
In the following experiments, we focus on spreadsheets specifically and hence only include \textit{functional} constraints, constraints that can be expressed as formulas. We ignoring structural constraints such as \textit{ALLDIFFERENT}, \textit{FOREIGNKEY}, \textit{PERMUTATION} and \textit{ORDERED}. All constraints are stored in their canonical form. 5 spreadsheets contained rows or columns that were entirely identical, and any intended constraint involving one such vector was also added for the equivalent vectors. \samuel{They are usually not copied}

Section~\ref{sec:which_cons} discusses the 33 constraints currently supported by \sname. Some of the intended constraints are currently outside of the scope of \sname however, more specifically nested mathematical or nested logical formulas. We denote by \textbf{supported} constraints the subset of intended constraints that the system can find in theory.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\linewidth]{figures/scatter_plots.pdf}
  \caption{Loglog plots of the experiments with synthetic (random) data. In plot~$a$ the number of vectors is increased; in plot~$b$ the size of the vectors is increased; and in plot~$c$ the number of blocks is increased (the block size is decreased).}
  \label{fig:runtime_analysis}
\end{figure}

@@

Our method uses dependencies to increase efficiency by finding constraints incrementally. To illustrate the effect this can have, we ran the benchmarks using FOREIGNKEY instead of ALLDIFFERENT as base constraint to find conditional aggregates such as SUMIF. This is a stronger assumption, and a too strong in practice as not all keys have to be used in the conditions. However, the total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data). This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.\sam{Check numbers} % (0.83 + 1.03 + 0.99) = 2.85 / (1.34 + 1.46 + 0.46) = 3.3  / (0.99 + 0.98 + 0.44) = 2.41

@@

% The intended constraints not discovered correspond to constraints that are not supported by the system (supported recall is 0.99), except for one intended \textit{SUMIF} constraint.
% The supported recall value mostly confirms that our bias (signature) is not too restrictive and is able to find constraints for supported constraint templates.
% By changing the signature of \textit{SUMIF} to allow \sbl{pk}
